### 2021-01-12 paper/main.tex
Zhao comments, math, typesetting

% to-do
% -----
% - Address the META over-fitting point that if you do lots of CVs and $n$ is small, you are probably still over-fitting. For example, the $p=73$ result is probably an over-fit. SV has an example in the literature to be cited. Goes in the discussion. But maybe also gets hinted in the introduction.
% - Incorporate remaining Zhao comments
% - Are we clear that the feature weighting is a kind of regularization? I (DWH) think it IS because it affects how the L2 on beta affects different parameters. SV says that maybe we should get explicit about the regularization points, and when the regularization is implicit or hidden.
% - In uncertainties, we talk about Bayesian. Should we also talk about frequentist and/or information theory?
% - Audit for use of \, as the multiply operator? Fix the kerning around the ^\top items?
% - HOGG: List clearly the assumptions underlying CVMSE and jackknife. They make assumptions about independence of the data points and also something maybe about per-point information content.
% - SOLEDAD: address constraints may not be satisfiable

@@

In some contexts you have strong beliefs about the noise affecting your measurements. In other cases you don't.
In some cases you have strong reasons to use a particular basis.
In other cases you don't.
The differences in these beliefs and the differences in your objectives will change what methods you choose and how you use and analyze them.
We'll try to be useful to you no matter where you're at.

@@

The idea of least-squares fitting is that the ``best'' values of the parameters $\beta_j$ are the values that minimize the sum of squares of the differences between the data and the linear combination of features:
\begin{equation}\label{eq:opt1}
    \hat{\beta} = \arg\min_\beta \|Y - X\,\beta\|_2^2
    ~,
\end{equation}
where $\hat{\beta}$ is the $p$-vector (column vector) of the $p$ best-fit values $\hat{\beta}_j$ of the parameters $\beta_j$, and $\|q\|_2^2$ denotes the squared L2-norm, or the \emph{sum of squares} of the components of $q$
\begin{equation}
    \|q\|_2^2 \equiv q^\top\,q
    ~.
\end{equation}
This optimization objective \eqref{eq:opt1} is convex and the optimization problem has a solution in closed form\footnote{
Note that $\nabla_{\beta} \|Y-X\beta\|^2 = 2X^\top(Y-X\beta)$ so critical points are such that $X^\top Y= X^\top X\beta$. Since the objective is convex all these critical points are global minima. 
} as long as the number of parameters (and features) $p$ is less than the number of training points $n$ (and the matrix $X^\top X$ is invertible):
\begin{equation}
    \hat{\beta} = (X^\top\,X)^{-1}\,X^\top\,Y
    ~.
    \label{OLS-under}
\end{equation}
We will treat the case in which the matrix is not invertible below.
When the investigator knows uncertainties on the training data $y_i$, this expression will change a bit; we begin to discuss that in the next \sectionname.

But recall our setting:
We are using the linear fit to interpolate the data, or de-noise the data, or predict new data.
In these contexts, we don't care about the parameter vector $\hat{\beta}$ itself.
We care only about the predictions at a set of new ``test'' locations $t_\ast$, which will usually be different from the training locations $t_j$.
From the new test times $t_\ast$ we create the test feature matrix $X_\ast$ (by the same feature embedding functions $g_j(t)$).
The prediction $\hat{Y}_\ast$ for the $y$ values at the test locations $t_\ast$ becomes
\begin{equation}\label{eq:OLS}
    \hat{Y}_\ast = X_\ast\,(X^\top\,X)^{-1}\,X^\top\,Y
    ~.
\end{equation}
Examples of OLS applied to some toy data are shown in \figurename~\ref{fig:ols1}.
This form \eqref{eq:OLS} of the prediction of new test data is called ordinary least squares (OLS). It has many good properties, some of which are encoded in the Gauss--Markov theorem.
In particular, if the noise contributions in the model given in equation \eqref{eq.model} are uncorrelated, have zero mean, and equal variances for $i=1,\ldots, n$ then the Gauss--Markov theorem states that the OLS estimator has the lowest variance within the class of unbiased estimators that are linear in $Y$ (see, for example, \citealt{esl}, Ch 3).

\begin{figure}[t]
    \begin{mdframed}
    \includegraphics[width=\figurewidth]{paper/OLS-under.pdf}
    \caption{Ordinary least-squares (OLS) fits (continuous lines) to a set of example data points (black dots). Fits are shown for different values of the number of basis functions $p$; there are $n=23$ data points. Here (and in all the \figurename s to follow) we are using the Fourier basis functions in \eqref{eq:basis} and shown in \figurename~\ref{fig:basis}. The data $y_i$ were generated using a function that does not reside in the function space spanned by the basis. Here the $X_\ast$ matrices used in the predictions $\hat{Y}_\ast$ were generated from a fine grid of locations in the location coordinate $t$; the plots of the fine grid of predictions are the continuous lines. The fits with larger $p$ have more flexibility to fit the data than the fit with $p=3$, but the fit at the highest $p$ shows evidence of over-fitting.}
    \label{fig:ols1}
    \end{mdframed}
\end{figure}

@@

Although the OLS prediction \eqref{eq:OLS} is affine invariant with respect to $p$-dimensional transformations, it is \emph{not} affine invariant with respect to $n$-dimensional transformations, such as a re-weighting of the input data. Indeed, if you know weights or inverse variances for your data points $y_i$, conceptually you can put them into a weight matrix $C^{-1}$ (written this way to emphasize that reweighting is usually inverse-variance weighting) and write
\begin{equation}\label{eq:wls}
    \hat{Y}_\ast = X_\ast\,(X^\top\,C^{-1}\,X)^{-1}\,X^\top\,C^{-1}\,Y
    ~,
\end{equation}
where the weight matrix $C^{-1}$ is $n\times n$ (and often diagonal in standard applications). The weight matrix $C^{-1}$ is sometimes called the information tensor (or information matrix) and its inverse $C$ is often called the covariance matrix or the noise variance tensor.

This form \eqref{eq:wls} of least squares is called \emph{weighted least squares} (WLS) because of the data weighting (not to be confused with \emph{feature} weighting, to appear below).
It is also called chi-squared fitting because it optimizes the scalar objective commonly called chi-squared:
\begin{equation}
    \chi^2 = (Y - X\,\beta)^\top\,C^{-1}\,(Y - X\,\beta)
    ~.
\end{equation}
(If that isn't obviously chi-squared to you, recall that $C^{-1}$ is often diagonal and has the inverses of the squares of the data uncertainties on that diagonal.)
In \figurename~\ref{fig:wls} a comparison of OLS and WLS is shown, for a case of non-trivial data weights, where the data weights are set to be the inverse squares of individual data-point uncertainties.
\begin{figure}[t]
    \begin{mdframed}
    \includegraphics[width=\figurewidth]{paper/WLS.pdf}
    \caption{Comparison of ordinary least-squares (OLS) and weighted least-squares (WLS) fits (continuous lines) to the example data (black dots). In order to illustrate the differences, we assigned non-trivial error bars to the data. The error bars are ignored in the OLS fit, but in the WLS fit, the weight matrix $C^{-1}$ is diagonal with the diagonal entries set to the inverses of the squares of those error bars. The WLS fit ``pays less attention to'' the points on the left with the largest error bars.}
    \label{fig:wls}
    \end{mdframed}
\end{figure}

@@

It is common to include a regularization that discourages the fit from making use of large amplitudes $\beta_j$.
There are many options, but the simplest is ridge regression (or Tikhonov regularization or L2 regularization), which (in the form that doesn't have data weights) looks like
\begin{equation} \label{eq:ridge}
    \hat{\beta} = \arg\min_\beta \|Y - X\,\beta\|_2^2 + \lambda\,\|\beta\|_2^2
    ~,
\end{equation}
where $\lambda>0$ is a regularization parameter that penalizes large values for elements $\beta_j$ of the parameter vector.
This optimization is also convex.
The ridge-regularized prediction for new data looks like
\begin{equation} \label{eq:ridge_sol}
    \hat{Y}_\ast = X_\ast\,(X^\top\,X + \lambda\,I)^{-1}\,X^\top\,Y
    ~,
\end{equation}
where $I$ is the $p\times p$ identity.
In the language of Bayesian inference, $\lambda$ can be seen as the inverse of a prior variance for the parameters $\beta_j$.
The salutary effect of the ridge regularization is shown in \figurename~\ref{fig:ridge}.
\begin{figure}[t]
    \begin{mdframed}
    \includegraphics[width=\figurewidth]{paper/ridge.pdf}
    \caption{Comparison of ordinary least-squares (OLS) and ridge-regression fits (continuous lines) to the example data (black dots). In order to illustrate the differences, we chose the $p=21$-parameter fit, which shows evidence of over-fitting at the edges of the fit range. The regularized fit looks more sensible, though it fits the individual data points less precisely. The choice of regularization parameter $\lambda$ matters; here we used $\lambda=0.1$ (chosen arbitrarily).}
    \label{fig:ridge}
    \end{mdframed}
\end{figure}

@@

You can combine both the point weighting from WLS and the generalized ridge regression into a weighted ridge that looks like
\begin{equation}\label{eq:foo}
    \hat{Y}_\ast = X_\ast\,(X^\top\,C^{-1}\,X + \Lambda)^{-1}\,X^\top\,C^{-1}\,Y
    ~.
\end{equation}
This form has good properties for many real-world physics applications, where data-point error bars are often known, and functions are often expected to be smooth.
We'll discuss this form \eqref{eq:foo} more below in \sectionname~\ref{sec:fwols},
but briefly we can say here that it appears in Bayesian inference contexts where the data points $y_i$ are treated as having Gaussian noise associated with them (with known variance tensor or covariance matrix $C$) and there is a Gaussian prior on the parameter vector $\beta$ (with known variance $\Lambda^{-1}$).
We have discussed that model elsewhere (\citealt{products}).
It is also useful sometimes to think of the ``units'' or dimensions of the quantities in \eqref{eq:foo}: In the Bayesian setting, the units of $C$ would be the square of the units of $Y$ and the units of $\Lambda^{-1}$ would be the square of the units of the ratio $Y/X$ (the square of the units of $\beta$).

@@

In the over-parameterized case ($p>n$), the prediction looks like
\begin{equation}\label{eq:OLS2}
    \hat{Y}_\ast = X_\ast\,X^\top\,(X\,X^\top)^{-1}\,Y
    ~,
\end{equation}
provided that $X\,X^\top$ is invertible (which will usually be the case).
Like \eqref{eq:OLS}, this prediction is \emph{also} called ordinary least squares (OLS), or sometimes ``min-norm least-squares'' to emphasize the point that it is making the minimum-norm choice of $\beta$ among many degenerate solutions.
Examples of OLS fits in the over-parameterized regime are shown in \figurename~\ref{fig:ols2}.
Note that the fits with different numbers of parameters $p$ lead to very different predictions, but they all go through the data exactly.
\begin{figure}[t]
    \begin{mdframed}
    \includegraphics[width=\figurewidth]{paper/OLS-over.pdf}
    \caption{Ordinary least-squares (OLS) or min-norm least-squares fits (continuous lines) to a set of example data points (black dots), but now for a few over-parameterized cases. There are $n=23$ data points. As in \figurename~\ref{fig:ols1}, the data points $y_i$ were generated using a function that does not precisely reside in the function space spanned by the basis. The three fits are very different, but they all go through all the data points exactly. As $p$ gets large, the fit function approaches $y=0$ almost everywhere; this is a consequence of Plancharel's Theorem (see text).}
    \label{fig:ols2}
    \end{mdframed}
\end{figure}

@@

The two equations for OLS---\eqref{eq:OLS} and \eqref{eq:OLS2}---can be unified into one equation (and also generalized to handle non-invertible matrices $X^\top X$ and $X\,X^\top$) if we define the pseudo-inverse $X^\dagger$:
\begin{equation}\label{eq:OLS3}
    \hat{Y}_\ast = X_\ast\,X^\dagger\,Y
    ~.
\end{equation}
The pseudo-inverse of a diagonal matrix is defined as the diagonal matrix made by inverting the non-zero diagonal entries. And for any non-diagonal (or any non-square) matrix $X$ the pseudo-inverse is defined by taking the singular-value decomposition (SVD) of $X$, $X=U\,S\,V$, with $S$ diagonal and $U,V$ orthogonal, then $X^\dagger \equiv V^\top\,S^\dagger\,U^\top$.

This pseudo-inverse form \eqref{eq:OLS3} of OLS is extremely general: It works for both the $p\le n$ and $p>n$ cases, and it works when the $X\,X^\top$ or $X^\top\,X$ matrices are not invertible.
There can be significant numerical issues with implementing the pseudo-inverse; we comment on those in \sectionname~\ref{sec:implementation}.

@@

The OLS prediction for $p>n$ is \emph{not} affine invariant with respect to $p$-dimensional rotations or rescalings.
That is, rotations and scalings in the feature space \emph{will} affect predictions.
It behooves us to re-scale the features (the $p$ $n$-vectors of the feature matrix $X$) in a sensible way, like for instance, to encourage the fit to use low-frequency features more than high-frequency features (\citealt{xie2020weighted, bah2016sample, rauhut2016interpolation}).
We can encode these feature weights in a $p\times p$ diagonal weight matrix $\Lambda^{-1}$ and the prediction becomes
\begin{equation} \label{eq:weighted_sol}
    \hat{Y}_\ast = X_\ast\,\Lambda^{-1}\,X^\top\,(X\,\Lambda^{-1}\,X^\top)^{-1}\,Y
\end{equation}
This can be seen as the prediction for new data resulting from the following optimization:
\begin{equation} \label{eq:weighted}
    \hat{\beta} = \arg\min_\beta \|\Lambda^{1/2}\,\beta\|_2^2 ~~\mbox{subject to}~~ Y = X\,\beta
    ~,
\end{equation}
which---in analogy to the optimizations \eqref{eq:opt2} and \eqref{eq:opt3}---can also be written as
\begin{equation}
    \hat{\beta} = \lim_{\lambda\rightarrow 0^+}\left[\arg\min_\beta \|Y - X\,\beta\|_2^2 + \lambda\,\|\Lambda^{1/2}\beta\|_2^2\right]
    ~.
\end{equation}
This optimization\footnote{If you are a physicist and you don't like the mathematical $\|\cdot\|_2^2$ notation---if you prefer to look at quantities that are obvious scalar forms---then recall that $\|\Lambda^{1/2}\,\beta\|_2^2 = \beta^\top\Lambda\,\beta$.
That right-hand-side object is a gauge-invariant object so long as $\beta$ and $\Lambda$ are also gauge-invariant objects themselves.
Yes, data analysis can have this kind of geometric structure!}
penalizes more strongly the parameters $\beta_j$ corresponding to features $g_j(t)$ with larger values of $[\Lambda^{1/2}]_{jj}$.

@@

In the most general case, you have not just a set of feature weights $\Lambda^{-1}$, you also have a set of data-point weights $C^{-1}$ (which, in standard settings, would be the inverses of the variances of the noise affecting the data points $y_i$).
When you put these all together, the feature-weighted, data-weighted least squares predictions are given by either of these two equivalent expressions:
\begin{align}\label{eq:LambdaC1}
    \hat{Y} &= X_\ast\,(X^\top\,C^{-1}\,X + \Lambda)^{-1}\,X^\top\,C^{-1}\,Y
    \\ \label{eq:LambdaC2}
    \hat{Y} &= X_\ast\,\Lambda^{-1}\,X^\top\,(X\,\Lambda^{-1}\,X^\top + C)^{-1}\,Y
    ~.
\end{align}
We discuss the equivalence of these below in \appendixname~\ref{app:math}. These expressions appear in Bayesian-inference contexts (see, for example \citealt{products}) when $\Lambda^{-1}$ is the variance of the prior on the parameter vector $\beta$ and $C$ is the variance of the noise on the data vector $Y$. In those contexts, the units of $C$ would be the units of $Y$ squared, and the units of $\Lambda^{-1}$ would be the square of the ratio of the units of $Y$ to the units of $X$ (the square of the units of $\beta$). These weighted-feature, weighted-data least-square forms are good because (if the weightings are chosen appropriately) they lead to smooth solutions that are not required to pass precisely through every data point. This is appropriate in common, real situations in which data are noisy and the world is smooth. Which form you choose, between \eqref{eq:LambdaC1} and \eqref{eq:LambdaC2}, depends on a few things, but primarily $p/n$. If $p<n$ it's both faster and more stable to use \eqref{eq:LambdaC1}; if $p>n$ it's faster and more stable to use \eqref{eq:LambdaC2}. We show a toy example of a fit with both feature weights $\Lambda^{-1}$ and data weights $C^{-1}$ in \figurename~\ref{fig:fwwls}.
\begin{figure}[t]
    \begin{mdframed}
    \includegraphics[width=\figurewidth]{paper/weighted-WLS.pdf}
    \caption{Comparison of the feature-weighted OLS shown in \figurename~\ref{fig:fwols} to the same but also including data weights $C^{-1}$, as in equations \eqref{eq:LambdaC1} and \eqref{eq:LambdaC2}.
    Similarly to \figurename~\ref{fig:wls}, the data weights on the diagonal of $C^{-1}$ are the inverses of the squares of the uncertainties shown as vertical error bars. The feature weights encourage the prediction to be smooth; the data weights permit it to be even smoother because they permit the prediction to miss the training data.
    Prediction results now depend on the specific amplitude or prefactor multiplying the feature weights; a particular value of $0.07\,f^2(\omega)$ was chosen for the diagonal of the feature weight matrix $\Lambda^{-1}$ for this demonstration.}
    \label{fig:fwwls}
    \end{mdframed}
\end{figure}

@@

Since it is an overloaded concept, we pause here to make a few comments about ``over-fitting''.... HOGG

@@

HOGG: INSERT POINT ABOUT OVER-CHOOSING YOUR HYPER-PARAMETERS: Over-fitting to the validation operator.

>>>>

% to-do
% -----
% - To be submitted to EPJ Data Science https://epjdatascience.springeropen.com/about
% - Audit for use of \, as the multiply operator?
% - HOGG: List clearly the assumptions underlying CVMSE and jackknife. They make assumptions about independence of the data points and also something maybe about per-point information content.
% - SOLEDAD: address the point that some constraints may not be satisfiable
% - Soledad: Once appendix is done, point down to it from the text at relevant places.

@@

In some contexts you have strong beliefs about the noise affecting your measurements. In other cases you don't.
In some cases you have strong reasons to use a particular basis.
In other cases you don't.
The differences in these beliefs and the differences in your objectives will change what methods you choose and how you use and analyze them.
We'll try to be useful to you no matter where you're at.
This document does not deliver new research results on the mathematics or statistics of regression.
It is novel only in that it makes very specific the connection between regularized linear regression and Gaussian processes.

@@

The idea of least-squares fitting is that the ``best'' values of the parameters $\beta_j$ are the values that minimize the sum of squares of the differences between the data and the linear combination of features:
\begin{equation}\label{eq:opt1}
    \hat{\beta} = \arg\min_\beta \|Y - X\,\beta\|_2^2
    ~,
\end{equation}
where $\hat{\beta}$ is the $p$-vector (column vector) of the $p$ best-fit values $\hat{\beta}_j$ of the parameters $\beta_j$, and $\|q\|_2^2$ denotes the squared L2-norm, or the \emph{sum of squares} of the components of $q$
\begin{equation}
    \|q\|_2^2 \equiv q^\top q
    ~.
\end{equation}
This optimization objective \eqref{eq:opt1} is convex and the optimization problem has a solution in closed form\footnote{
Note that $\nabla_{\beta} \|Y-X\beta\|^2 = 2X^\top(Y-X\beta)$ so critical points are such that $X^\top Y= X^\top X\beta$. Since the objective is convex all these critical points are global minima. 
} as long as the number of parameters (and features) $p$ is less than the number of training points $n$ (and the matrix $X^\top X$ is invertible):
\begin{equation}
    \hat{\beta} = (X^\top X)^{-1}\,X^\top Y
    ~.
    \label{OLS-under}
\end{equation}
We will treat the case in which the matrix is not invertible below.
When the investigator knows uncertainties on the training data $y_i$, this expression will change a bit; we begin to discuss that in the next \sectionname.

But recall our setting:
We are using the linear fit to interpolate the data, or de-noise the data, or predict new data.
In these contexts, we don't care about the parameter vector $\hat{\beta}$ itself.
We care only about the predictions at a set of new ``test'' locations $t_\ast$, which will usually be different from the training locations $t_j$.
From the new test times $t_\ast$ we create the test feature matrix $X_\ast$ (by the same feature embedding functions $g_j(t)$).
The prediction $\hat{Y}_\ast$ for the $y$ values at the test locations $t_\ast$ becomes
\begin{equation}\label{eq:OLS}
    \hat{Y}_\ast = X_\ast\,(X^\top X)^{-1}\,X^\top Y
    ~.
\end{equation}
Examples of OLS applied to some toy data\footnote{The algorithm by which the toy data were made---and indeed all of the code executed to make the figures for this \documentname---is available online at \url{https://github.com/davidwhogg/FlexibleLinearModels}.} are shown in \figurename~\ref{fig:ols1}.
This form \eqref{eq:OLS} of the prediction of new test data is called ordinary least squares (OLS). It has many good properties, some of which are encoded in the Gauss--Markov theorem.
In particular, if the noise contributions in the model given in equation \eqref{eq.model} are uncorrelated, have zero mean, and equal variances for $i=1,\ldots, n$ then the Gauss--Markov theorem states that the OLS estimator has the lowest variance within the class of unbiased estimators that are linear in $Y$ (see, for example, \citealt{esl}, Ch 3).
\begin{figure}[t]
    \begin{mdframed}
    \includegraphics[width=\figurewidth]{paper/OLS-under.pdf}
    \caption{Ordinary least-squares (OLS) fits (continuous lines) to a set of example data points (black dots). Fits are shown for different values of the number of basis functions $p$; there are $n=23$ data points. Here (and in all the \figurename s to follow) we are using the Fourier basis functions in \eqref{eq:basis} and shown in \figurename~\ref{fig:basis}. The data $y_i$ were generated using a function that does not reside in the function space spanned by the basis. Here the $X_\ast$ matrices used in the predictions $\hat{Y}_\ast$ were generated from a fine grid of locations in the location coordinate $t$; the plots of the fine grid of predictions are the continuous lines. The fits with larger $p$ have more flexibility to fit the data than the fit with $p=3$, but the fit at the highest $p$ shows evidence of over-fitting.}
    \label{fig:ols1}
    \end{mdframed}
\end{figure}

@@

Although the OLS prediction \eqref{eq:OLS} is affine invariant with respect to $p$-dimensional transformations, it is \emph{not} affine invariant with respect to $n$-dimensional transformations, such as a re-weighting of the input data. Indeed, if you know weights or inverse variances for your data points, conceptually you can put them into a weight matrix $C^{-1}$ (written this way to emphasize that reweighting is usually inverse-variance weighting) and write
\begin{equation}\label{eq:wls}
    \hat{Y}_\ast = X_\ast\,(X^\top C^{-1}\,X)^{-1}\,X^\top C^{-1}\,Y
    ~,
\end{equation}
where the weight matrix $C^{-1}$ is $n\times n$ (and often diagonal in standard applications). The weight matrix $C^{-1}$ is sometimes called the information tensor (or information matrix) and its inverse $C$ is often called the covariance matrix or the noise variance tensor.

This form \eqref{eq:wls} of least squares is called \emph{weighted least squares} (WLS) because of the data weighting (not to be confused with \emph{feature} weighting, to appear below).
It is also called chi-squared fitting because it optimizes the scalar objective commonly called chi-squared:
\begin{equation}
    \chi^2 = (Y - X\,\beta)^\top C^{-1}\,(Y - X\,\beta)
    ~.
\end{equation}
(If that isn't obviously chi-squared to you, recall that $C^{-1}$ is often diagonal and has the inverses of the squares of the data uncertainties on that diagonal.)
In \figurename~\ref{fig:wls} a comparison of OLS and WLS is shown, for a case of non-trivial data weights, where the data weights are set to be the inverse squares of individual data-point uncertainties.
\begin{figure}[t]
    \begin{mdframed}
    \includegraphics[width=\figurewidth]{paper/WLS.pdf}
    \caption{Comparison of ordinary least-squares (OLS) and weighted least-squares (WLS) fits (continuous lines) to the example data (black dots). In order to illustrate the differences, we assigned non-trivial error bars to the data. The error bars are ignored in the OLS fit, but in the WLS fit, the weight matrix $C^{-1}$ is diagonal with the diagonal entries set to the inverses of the squares of those error bars. The WLS fit ``pays less attention to'' the points on the left with the largest error bars.}
    \label{fig:wls}
    \end{mdframed}
\end{figure}

@@

It is common to include a regularization that discourages the fit from making use of large amplitudes $\beta_j$.
There are many options, but the simplest is ridge regression (or Tikhonov regularization or L2 regularization), which (in the form that doesn't have data weights) looks like
\begin{equation} \label{eq:ridge}
    \hat{\beta} = \arg\min_\beta \|Y - X\,\beta\|_2^2 + \lambda\,\|\beta\|_2^2
    ~,
\end{equation}
where $\lambda>0$ is a regularization parameter that penalizes large values for elements $\beta_j$ of the parameter vector.
This optimization is also convex.
The ridge-regularized prediction for new data looks like
\begin{equation} \label{eq:ridge_sol}
    \hat{Y}_\ast = X_\ast\,(X^\top X + \lambda\,I)^{-1}\,X^\top Y
    ~,
\end{equation}
where $I$ is the $p\times p$ identity.
In the language of Bayesian inference, $\lambda$ can be seen as the inverse of a prior variance for the parameters $\beta_j$.
The salutary effect of the ridge regularization is shown in \figurename~\ref{fig:ridge}.
\begin{figure}[t]
    \begin{mdframed}
    \includegraphics[width=\figurewidth]{paper/ridge.pdf}
    \caption{Comparison of ordinary least-squares (OLS) and ridge-regression fits (continuous lines) to the example data (black dots). In order to illustrate the differences, we chose the $p=21$-parameter fit, which shows evidence of over-fitting at the edges of the fit range. The regularized fit looks more sensible, though it fits the individual data points less precisely. The choice of regularization parameter $\lambda$ matters; here we used $\lambda=0.1$ (chosen arbitrarily).}
    \label{fig:ridge}
    \end{mdframed}
\end{figure}

@@

You can combine both the point weighting from WLS and the generalized ridge regression into a weighted ridge that looks like
\begin{equation}\label{eq:foo}
    \hat{Y}_\ast = X_\ast\,(X^\top C^{-1}\,X + \Lambda)^{-1}\,X^\top C^{-1}\,Y
    ~.
\end{equation}
This form has good properties for many real-world physics applications, where data-point error bars are often known, and functions are often expected to be smooth.
We'll discuss this form \eqref{eq:foo} more below in \sectionname~\ref{sec:fwols},
but briefly we can say here that it appears in Bayesian inference contexts where the data points $y_i$ are treated as having Gaussian noise associated with them (with known variance tensor or covariance matrix $C$) and there is a Gaussian prior on the parameter vector $\beta$ (with known variance $\Lambda^{-1}$).
We have discussed that model elsewhere (\citealt{products}).\footnote{In our previous discussion of this product of Gaussians, the notation differs. What's called $\Lambda^{-1}$ here is called $\Lambda$ in \citet{products}.}
It is also useful sometimes to think of the ``units'' or dimensions of the quantities in \eqref{eq:foo}: In the Bayesian setting, the units of $C$ would be the square of the units of $Y$ and the units of $\Lambda^{-1}$ would be the square of the units of the ratio $Y/X$ (the square of the units of $\beta$).

@@

In the over-parameterized case ($p>n$), the prediction looks like
\begin{equation}\label{eq:OLS2}
    \hat{Y}_\ast = X_\ast\,X^\top (X\,X^\top)^{-1}\,Y
    ~,
\end{equation}
provided that $X\,X^\top$ is invertible (which will usually be the case).
Like \eqref{eq:OLS}, this prediction is \emph{also} called ordinary least squares (OLS), or sometimes ``min-norm least-squares'' to emphasize the point that it is making the minimum-norm choice of $\beta$ among many degenerate solutions.
Examples of OLS fits in the over-parameterized regime are shown in \figurename~\ref{fig:ols2}.
Note that the fits with different numbers of parameters $p$ lead to very different predictions, but they all go through the data exactly.
\begin{figure}[t]
    \begin{mdframed}
    \includegraphics[width=\figurewidth]{paper/OLS-over.pdf}
    \caption{Ordinary least-squares (OLS) or min-norm least-squares fits (continuous lines) to a set of example data points (black dots), but now for a few over-parameterized cases. There are $n=23$ data points. As in \figurename~\ref{fig:ols1}, the data points $y_i$ were generated using a function that does not precisely reside in the function space spanned by the basis. The three fits are very different, but they all go through all the data points exactly. As $p$ gets large, the fit function approaches $y=0$ almost everywhere; this is a consequence of Plancharel's Theorem (see text).}
    \label{fig:ols2}
    \end{mdframed}
\end{figure}

@@

The two equations for OLS---\eqref{eq:OLS} and \eqref{eq:OLS2}---can be unified into one equation (and also generalized to handle non-invertible matrices $X^\top X$ and $X\,X^\top$) if we define the pseudo-inverse $X^\dagger$:
\begin{equation}\label{eq:OLS3}
    \hat{Y}_\ast = X_\ast\,X^\dagger\,Y
    ~.
\end{equation}
The pseudo-inverse of a diagonal matrix is defined as the diagonal matrix made by inverting the non-zero diagonal entries. And for any non-diagonal (or any non-square) matrix $X$ the pseudo-inverse is defined by taking the singular-value decomposition (SVD) of $X$, $X=U\,S\,V$, with $S$ diagonal and $U,V$ orthogonal, then $X^\dagger \equiv V^\top S^\dagger\,U^\top$.

This pseudo-inverse form \eqref{eq:OLS3} of OLS is extremely general: It works for both the $p\le n$ and $p>n$ cases, and it works when the $X\,X^\top$ or $X^\top X$ matrices are not invertible.
There can be significant numerical issues with implementing the pseudo-inverse; we comment on those in \sectionname~\ref{sec:implementation}.

@@

The OLS prediction for $p>n$ is \emph{not} affine invariant with respect to $p$-dimensional rotations or rescalings.
That is, rotations and scalings in the feature space \emph{will} affect predictions.
It behooves us to re-scale the features (the $p$ $n$-vectors of the feature matrix $X$) in a sensible way, like for instance, to encourage the fit to use low-frequency features more than high-frequency features (\citealt{xie2020weighted, bah2016sample, rauhut2016interpolation}).
We can encode these feature weights in a $p\times p$ diagonal weight matrix $\Lambda^{-1}$ and the prediction becomes
\begin{equation} \label{eq:weighted_sol}
    \hat{Y}_\ast = X_\ast\,\Lambda^{-1}\,X^\top (X\,\Lambda^{-1}\,X^\top)^{-1}\,Y
\end{equation}
This can be seen as the prediction for new data resulting from the following optimization (again, assuming the data can be interpolated):
\begin{equation} \label{eq:weighted}
    \hat{\beta} = \arg\min_\beta \|\Lambda^{1/2}\,\beta\|_2^2 ~~\mbox{subject to}~~ Y = X\,\beta
    ~,
\end{equation}
which---in analogy to the optimizations \eqref{eq:opt2} and \eqref{eq:opt3}---can also be written as
\begin{equation}
    \hat{\beta} = \lim_{\lambda\rightarrow 0^+}\left[\arg\min_\beta \|Y - X\,\beta\|_2^2 + \lambda\,\|\Lambda^{1/2}\,\beta\|_2^2\right]
    ~.
\end{equation}
This optimization\footnote{If you are a physicist and you don't like the mathematical $\|\cdot\|_2^2$ notation---if you prefer to look at quantities that are obvious scalar forms---then recall that $\|\Lambda^{1/2}\,\beta\|_2^2 = \beta^\top\Lambda\,\beta$.
That right-hand-side object is a gauge-invariant object so long as $\beta$ and $\Lambda$ are also gauge-invariant objects themselves.
Yes, data analysis can have this kind of geometric structure!}
penalizes more strongly the parameters $\beta_j$ corresponding to features $g_j(t)$ with larger values of $[\Lambda^{1/2}]_{jj}$.

@@

In the most general case, you have not just a set of feature weights $\Lambda^{-1}$, you also have a set of data-point weights $C^{-1}$ (which, in standard settings, would be the inverses of the variances of the noise affecting the data points $y_i$).
When you put these all together, the feature-weighted, data-weighted least squares predictions are given by either of these two equivalent expressions:
\begin{align}\label{eq:LambdaC1}
    \hat{Y} &= X_\ast\,(X^\top C^{-1}\,X + \Lambda)^{-1}\,X^\top C^{-1}\,Y
    \\ \label{eq:LambdaC2}
    \hat{Y} &= X_\ast\,\Lambda^{-1}\,X^\top (X\,\Lambda^{-1}\,X^\top + C)^{-1}\,Y
    ~.
\end{align}
We discuss the equivalence of these below in \appendixname~\ref{app:math}. These expressions appear in Bayesian-inference contexts (see, for example, \citealt{products}) when $\Lambda^{-1}$ is the variance of the prior on the parameter vector $\beta$ and $C$ is the variance of the noise on the data vector $Y$.
They are solutions to the optimization
\begin{equation}
    \hat{\beta} = \arg\min_\beta \|C^{-1/2}\,(Y - X\,\beta)\|_2^2 + \|\Lambda^{1/2}\,\beta\|_2^2
    ~.
\end{equation}
The units of $C$ must be the units of $Y$ squared, and the units of $\Lambda^{-1}$ must be the square of the ratio of the units of $Y$ to the units of $X$ (the square of the units of $\beta$). These weighted-feature, weighted-data least-square forms \eqref{eq:LambdaC1} and \eqref{eq:LambdaC2} are good because (if the weightings are chosen appropriately) they lead to smooth solutions that are not required to pass precisely through every data point. This is appropriate in common, real situations in which data are noisy and the world is smooth. Which form you choose, between \eqref{eq:LambdaC1} and \eqref{eq:LambdaC2}, depends on a few things, but primarily $p/n$. If $p<n$ it's both faster and more stable to use \eqref{eq:LambdaC1}; if $p>n$ it's faster and more stable to use \eqref{eq:LambdaC2}. We show a toy example of a fit with both feature weights $\Lambda^{-1}$ and data weights $C^{-1}$ in \figurename~\ref{fig:fwwls}.
\begin{figure}[t]
    \begin{mdframed}
    \includegraphics[width=\figurewidth]{paper/weighted-WLS.pdf}
    \caption{Comparison of the feature-weighted OLS shown in \figurename~\ref{fig:fwols} to the same but also including data weights $C^{-1}$, as in equations \eqref{eq:LambdaC1} and \eqref{eq:LambdaC2}.
    Similarly to \figurename~\ref{fig:wls}, the data weights on the diagonal of $C^{-1}$ are the inverses of the squares of the uncertainties shown as vertical error bars. The feature weights encourage the prediction to be smooth; the data weights permit it to be even smoother because they permit the prediction to miss the training data.
    Prediction results now depend on the specific amplitude or prefactor multiplying the feature weights; a particular value of $0.07\,f^2(\omega)$ was chosen for the diagonal of the feature weight matrix $\Lambda^{-1}$ for this demonstration.}
    \label{fig:fwwls}
    \end{mdframed}
\end{figure}

@@

Sometimes the word ``over-fitting'' is used to describe models that are too flexible.
In our view, a model is properly called ``over-fit'' when the prediction of the regression on held-out data---what the statisticians call ``risk''---is bad.
So over-fitting happens not universally when the number of parameters is large, but instead when the number of parameters is comparable to the number of data points, and (also) the regularization is inappropriate.
\figurename~\ref{fig:cv} shows that models with both small and large numbers $p$ of parameters can make good predictions for held-out data, and that regularization can also protect a model from over-fitting, even when $p\approx n$.
We don't mean to imply that over-fitting is not a problem in regression; it can be!
Our recommendation is just that the problem of over-fitting be analyzed empirically through cross-validation, not through intuitive ideas about the number of parameters.

@@

There is one more point to make about over-fitting which is that, when your data set is small, it is possible to over-fit even your cross-validation.
This problem is beyond the scope of this \documentname; all we will say here is that it is not possible to make precise settings of many hyper-parameters through cross validation.
Because in cross validation you are making use of your data to set the properties of your regression, there are dangers of over-adapting your method to your data.
It is safer to have a fully independent validation data set, but this is rarely practical (and it doesn't completely protect you either \citealt{SOLEDAD}).
